proof-of-concept_mods

### 2023-10-12 combat_sandbox.Rmd
playing around with different combats' variance outputs, plots

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv)
#install.packages("devtools")
devtools::install_github("andy1764/ComBatFamily", build_vignettes = FALSE) #per discussion with Andrew, sent vignettes separately
library(ComBatFamily)
```

@@

simple combat with linear model for covariates
```{r}
cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
```

combat with age as smooth
```{r}
cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
```

>>>>

Playing around with different combat configs using the UKB data, which has been assigned to 3 simulated study sites (see `proof-of-concept_mods.Rmd`)

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv, gamlss)
#install.packages("devtools")
devtools::install_github("andy1764/ComBatFamily", build_vignettes = FALSE) #per discussion with Andrew, sent vignettes separately
library(ComBatFamily)
```

@@

simple combat with no covariate preservation
```{r}
cf <- comfam(rvol.df, batch)
cf$dat.combat
cf$dat.combat[, "sim.site"] <- batch
```

simple combat with linear model for covariates
```{r}
cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
cf.lm$dat.combat
```
plot
```{r}
dim(cf.lm$dat.combat)[1] == length(batch)
#add back site
cf.lm$dat.combat[, "sim.site"] <- batch

ggplot() + 
 #geom_histogram(aes(x=cf.lm$dat.combat$GMV, y=after_stat(density), fill=cf.lm$dat.combat$sim.site, color=cf.lm$dat.combat$sim.site), alpha = .2)+
 geom_density(aes(x=cf.lm$dat.combat$lh_Vol_superiorparietal, fill=cf.lm$dat.combat$sim.site, color=cf.lm$dat.combat$sim.site), alpha=.4) +
  geom_vline(aes(xintercept=cf.lm_summary$sd_below.lh_v_sp, color=cf.lm_summary$sim.site)) +
  geom_vline(aes(xintercept=cf.lm_summary$sd_above.lh_v_sp, color=cf.lm_summary$sim.site))
```

Try setting site A (which has even sex distribution) as ref

```{r}
cf.lm2 <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
cf.lm2$dat.combat
cf.lm2$dat.combat[, "sim.site"] <- batch
```

combat gam with age as smooth
```{r}
cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
cf.gam$dat.combat[, "sim.site"] <- batch
```

```{r}
cf.gam2 <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
cf.gam2$dat.combat[, "sim.site"] <- batch
```

combat gamlss!

```{r}
cf.gamlss <- comfam(rvol.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
              sigma.formula = ~ age_days + sexMale, 
              control = gamlss.control(trace = FALSE))
cf.gamlss$dat.combat[, "sim.site"] <- batch
```

```{r}
cf.gamlss2 <- comfam(rvol.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
              sigma.formula = ~ age_days + sexMale, 
              control = gamlss.control(trace = FALSE),
              ref.batch = "Site_A")
cf.gamlss2$dat.combat[, "sim.site"] <- batch
```

## Compare Variances

Randomly picked `lh_Vol_superiorparietal` to compare

Get variances
```{r}
#raw data
sum_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal),
            var.lh_v_sp = var(lh_Vol_superiorparietal))
#no covars
cf_summary <- cf$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple lm
cf.lm_summary <- cf.lm$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple lm with ref site
cf.lm2_summary <- cf.lm2$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam
cf.gam_summary <- cf.gam$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam with ref site
cf.gam2_summary <- cf.gam2$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam
cf.gamlss_summary <- cf.gamlss$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam with ref site
cf.gamlss2_summary <- cf.gamlss2$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
```
compile

```{r}
var_comp_df <- data.frame(sim.site = sim.site.list,
                          raw_data = sum_df$var.lh_v_sp,
                          cf = cf_summary$var.lh_v_sp,
                          cf.lm = cf.lm_summary$var.lh_v_sp,
                          cf.lm2 = cf.lm2_summary$var.lh_v_sp,
                          cf.gam = cf.gam_summary$var.lh_v_sp,
                          cf.gam2 = cf.gam2_summary$var.lh_v_sp,
                          cf.gamlss = cf.gamlss_summary$var.lh_v_sp,
                          cf.gamlss2 = cf.gamlss2_summary$var.lh_v_sp)

var_comp_long <- var_comp_df %>%
  pivot_longer(!sim.site, names_to="type", values_to = "variance")

ggplot(var_comp_long) +
  geom_col(aes(x=type, y=variance, fill=sim.site, color=sim.site), position = "dodge")

ggplot(var_comp_long) +
  geom_col(aes(x=type, y=variance), position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~sim.site)
```

### 2023-10-12 proof-of-concept_mods.Rmd
playing around with different combats' variance outputs, plots

```{r}

```

>>>>

## Simulate Sites
Dividing each subject into 3 sites (A, B, and C) with M:F ratios of 1:1, 1:4, and 4:1

```{r}
set.seed(1010101)
sim.site.list <- c("Site_A", "Site_B", "Site_C")
n_female <- table(ukb_df_age.filt$sex)["Female"]
n_male <- table(ukb_df_age.filt$sex)["Male"]

# Sample probabilities
male_prob <- c(0.33, 0.0825, 0.5875)
female_prob <- c(0.33, 0.5875, 0.0825)

# Separate samples for males and females
sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)

ukb_df_sites <- ukb_df_age.filt %>%
  mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))

#write out
write_csv(ukb_df_sites, file="data/ukb_CN_data_simsites.csv")
```

```{r}
table(ukb_df_sites$sim.site)
table(ukb_df_sites$sim.site, ukb_df_sites$sex)
sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
anova(lm(age_days ~ sim.site, ukb_df_sites))
```

Need to come up with cleaner way to do this...

```{r}
site_sum_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  summarise(proportion_male = mean(sex == "Male"),
            min_age = min(age_days), max_age = max(age_days), mean_age = mean(age_days), 
            var_gmv = var(GMV), var_wmv=var(WMV), var_sgmv=var(sGMV), var_ven = var(Ventricles),
            sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
```

Center and visualize distributions

```{r}
ukb_sites_sex.centered <- ukb_df_sites %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

```{r}
# Generate histograms (mean-centered w/in sex)
global_site_plots <- plot.hist.by.sex(ukb_sites_sex.centered, vol_list_global, facet_fac = "sim.site")

# Print or display the histograms
for (pheno in vol_list_global) {
  print(global_site_plots[[pheno]])
}
```

Look at uncentered distributions across sites
```{r}
ggplot(ukb_df_sites, aes(x=GMV)) + 
 geom_histogram(aes(y=after_stat(density)))+
 geom_density(alpha=.2) +
  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV))))

ggplot(ukb_df_sites, aes(x=GMV)) + 
 geom_histogram(aes(y=after_stat(density)))+
 geom_density(alpha=.2) +
  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV)))) +
  facet_wrap(~sim.site)
```

```{r}
centered.site_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()

centered.site_sum_df <- centered.site_df %>%
  group_by(sim.site) %>%
  summarise(sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))

ggplot() + 
 #geom_histogram(aes(x=centered.site_df$GMV, y=after_stat(density), fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha = .2)+
 geom_density(aes(x=centered.site_df$GMV, fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha=.4) +
  geom_vline(aes(xintercept=centered.site_sum_df$sd_below.gmv, color=centered.site_sum_df$sim.site)) +
  geom_vline(aes(xintercept=centered.site_sum_df$sd_above.gmv, color=centered.site_sum_df$sim.site))
```

## ComBat

And now, to combat the data across the simulated sites!
